### 2022-11-24 ms.rmd
New tree

---
title: ""
author: et al. 
date: "`r Sys.Date()`"
bibliography: ./bib/refs.bib
csl: ./bib/ecology-letters.csl
output: 
  bookdown::word_document2:
    toc: no
    toc_depth: 6
    number_sections: false
    reference_docx: ./bib/template.docx
  bookdown::html_document2:
    code_folding: hide
    number_sections: no
    toc: yes
    toc_depth: 6
    toc_float: yes
    bibliography: refs.bib
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, tidy = TRUE, fig.width = 10)
## numbers >= 10^5 will be denoted in scientific notation,		  ## numbers >= 10^5 will be denoted in scientific notation,
  ## and rounded to 2 digits		  ## and rounded to 2 digits
  options(digits = 2)

##################################
# Clean workspace
##################################
    rm(list=ls())

##################################
# Loading packages & Functions
##################################
    pacman::p_load(tidyverse,  metafor, flextable,  patchwork, ggtree, ape, phytools, viridis, MCMCglmm, brms, latex2exp, orchaRd, multcomp)

```

```{r, tree}
# Bring in the fly tree
tree <- read.tree("./phylogeny/drosi2.tre")

# Plot the tree
plot(tree, cex = 0.2)

# Ultrametric?
is.ultrametric(tree) # nope
tree2 <- force.ultrametric(tree, method = "extend")

# Grab out data
   data <- read.csv("./data/drosi_data.csv")

# functions
source("./R/func.R")
t <- tree_checks(tree2, data = data, dataCol = "species", type = "check")
data <- data %>% mutate(species = gsub(" ", "", species),
                        phylo = species,
                        obs = 1:n()) %>% filter(!species %in% t$Species_InData_But_NotTree) 
tree3 <- tree_checks(tree2, data = data, dataCol = "species", type = "prune")
tree_checks(tree3, data = data, dataCol = "species", type = "check")
# Grab phylo matrix
phy_vcv <- vcv(tree3, cor = FALSE)

```

```{r}

## Model(s)
   ctmax <- bf(ctmax | se(sqrt(ctmax_se)) ~ 1 + (1|q|gr(phylo, cov = phy_vcv)) + (1|species) + (1|obs)) + gaussian()
    cfts <- bf( cfts |  mi() ~ 1 + (1|q|gr(phylo, cov = phy_vcv)) + (1|species)) + gaussian()
   #tdt_i <- bf(tdt_i |  mi() ~ 1 + (1|q|gr(phylo = phy_vcv)) + (1|spp)) + gaussian()
   tdt_z <- bf(tdt_z |  mi() ~ 1 + (1|q|gr(phylo, cov = phy_vcv)) + (1|species)) + gaussian()

       mv_brms_model <- brms::brm(ctmax + tdt_z + cfts, 
                        warmup = 1000, iter = 4000, thin = 5, 
                        control = list(adapt_delta = 0.95, max_treedepth = 15), cores = 4,
                        data = data, data2 = list(phy_vcv = phy_vcv))

#### Try MCMCglmm which I expect will be MUCH faster. Deals with missing data using data augmentation

prior = list(R = list(V = diag(3), nu=1.002), 
             G = list(G1=list(V = diag(3), nu = 1.002)))

data$animal = data$phylo

mcmcglmm_model <- MCMCglmm::MCMCglmm(cbind(ctmax, cfts, tdt_z) ~ -1 + trait, random = ~us(trait):animal, rcov = ~us(trait):units ,
                                     pedigree=tree3,
                                     burnin=1000, nitt = 300000, thin = 5, family = rep("gaussian", 3), prior = prior, 
                                    pl = TRUE, pr = TRUE, data = data)
saveRDS(mcmcglmm_model, file = "./output/model/mcmcglmm_model")
```

## References