### 2022-11-26 ms.rmd
Added node labels = NULL and added a models with scaled variables to make it easier to check correlation at phlo level

---
title: ""
author: et al. 
date: "`r Sys.Date()`"
bibliography: ./bib/refs.bib
csl: ./bib/ecology-letters.csl
output: 
  bookdown::word_document2:
    toc: no
    toc_depth: 6
    number_sections: false
    reference_docx: ./bib/template.docx
  bookdown::html_document2:
    code_folding: hide
    number_sections: no
    toc: yes
    toc_depth: 6
    toc_float: yes
    bibliography: refs.bib
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, tidy = TRUE, fig.width = 10)
## numbers >= 10^5 will be denoted in scientific notation,		  ## numbers >= 10^5 will be denoted in scientific notation,
  ## and rounded to 2 digits		  ## and rounded to 2 digits
  options(digits = 2)

##################################
# Clean workspace
##################################
    rm(list=ls())

##################################
# Loading packages & Functions
##################################
    pacman::p_load(tidyverse,  metafor, flextable,  patchwork, ggtree, ape, phytools, viridis, MCMCglmm, brms, latex2exp, orchaRd, multcomp)

```

```{r, tree}
# Bring in the fly tree
tree <- read.tree("./phylogeny/drosi2.tre")

# Plot the tree
plot(tree, cex = 0.2)

# Ultrametric?
is.ultrametric(tree) # nope
tree2 <- force.ultrametric(tree, method = "extend")

# Grab out data
   data <- read.csv("./data/drosi_data.csv")

# functions
source("./R/func.R")
t <- tree_checks(tree2, data = data, dataCol = "species", type = "check")
data <- data %>% mutate(species = gsub(" ", "", species),
                        phylo = species,
                        obs = 1:n()) %>% filter(!species %in% t$Species_InData_But_NotTree) 
tree3 <- tree_checks(tree2, data = data, dataCol = "species", type = "prune")
tree_checks(tree3, data = data, dataCol = "species", type = "check")
# Grab phylo matrix
phy_vcv <- vcv(tree3, cor = FALSE)
tree3$node.label <- NULL

```

```{r}

#### Try MCMCglmm which I expect will be MUCH faster. Deals with missing data using data augmentation
rerun = FALSE
if(rerun){
  prior = list(R = list(V = diag(3), nu=1.002), 
             G = list(G1=list(V = diag(3), nu = 1.002)))

data$animal = data$phylo

mcmcglmm_model <- MCMCglmm::MCMCglmm(cbind(ctmax, cfts, tdt_z) ~ -1 + trait, random = ~us(trait):animal, rcov = ~us(trait):units ,pedigree=tree3,burnin=1000, nitt = 300000, thin = 5, family = rep("gaussian", 3), prior = prior, pl = TRUE, pr = TRUE, data = data)

# We'll also z-transofmr to make the coprrelation stuff easier to calculate. 
data <- data %>% mutate(z_ctmax = scale(ctmax),
                         z_cfts = scale(cfts),
                        z_tdt_z = scale(tdt_z))

mcmcglmm_model_z_trans <- MCMCglmm::MCMCglmm(cbind(z_ctmax, z_cfts, z_tdt_z) ~ -1 + trait, random = ~us(trait):animal, rcov = ~us(trait):units, pedigree=tree3,burnin=1000, nitt = 300000, thin = 5, family = rep("gaussian", 3), prior = prior, pl = TRUE, pr = TRUE, data = data)

saveRDS(mcmcglmm_model, file = "./output/model/mcmcglmm_model")
saveRDS(mcmcglmm_model_z_trans, file = "./output/model/mcmcglmm_model_z_trans")
} else{
  mcmcglmm_model <- readRDS("./output/model/mcmcglmm_model")
}

# Extract estimates
sol <- mcmcglmm_model$Sol
vcv <- mcmcglmm_model$VCV

# Extract the estmated / imputaed Z for each species, but drop internal ancestral nodes
z <- sol[, grep("traittdt_z.animal", colnames(sol))]
z <- z[,-grep("Node", colnames(z))]
head(z)

# Clean up names so they are just species names

```

## References